---
title: "Design Challenge 1"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r}
library(tidyverse)
library(ggrepel)
library(scales)
library(viridis)

census <- read.csv(file="census.csv", header=TRUE, sep=",")
View(census)

```

```{r}
#highest_income <- census %>% arrange(-Median_Household_Income_2015) %>% #select(Area_name, Median_Household_Income_2015, Med_HH_Income_Percent_of_State_Total_2015)


hi_cut <- census %>% filter(Median_Household_Income_2015 > 100000)

hi_percent_cut <- highest_income %>% filter(Med_HH_Income_Percent_of_State_Total_2015 > 140)

states <- census %>% filter(is.na(X2003.Rural.urban.Continuum.Code) & !FIPS %in% c(2105, 2195, 2198, 2230, 2275))
states

counties <- census %>% filter(!is.na(X2003.Rural.urban.Continuum.Code) & !FIPS %in% c(2105, 2195, 2198, 2230, 2275))
counties
```

```{r}
bar_income_per_state_avg <- ggplot() + geom_bar(data=hi_cut, aes(reorder(Area_name, Median_Household_Income_2015), Median_Household_Income_2015, fill=Med_HH_Income_Percent_of_State_Total_2015), stat="identity") +
  coord_flip() +
  labs(x='', y='Median household income in census area ($)', fill="Household income as\n% of state average", 
       title="Census areas with the highest median household income are almost all located in\nstates with high average incomes.", subtitle="\nTwo major exceptions are Williamson County, TN, a suburb of Nashville that boasts many of\nTennessee's richest, and Los Alamos County, NM, home to a well-paying national laboratory...\nand not much else.") +
  scale_y_continuous(labels = function(x) paste0("$", scales::comma(x))) +
  scale_fill_viridis(option="plasma", limits=c(140,240), breaks=seq(140,240,20), labels=function(x) paste0(x, "%"))+
  guides(fill = guide_colourbar(barheight = 8))
#+
 # guides(fill = guide_colourbar(labels=c(140,160,180,200,220,240)))

bar_income_per_state_avg

ggsave(filename="bar_income_per_state_avg.pdf", device = "pdf", plot = bar_income_per_state_avg, height = 5, width = 8.5, units = "in")
```

```{r}
hi_cut_minus_LA <- hi_cut %>% filter(Med_HH_Income_Percent_of_State_Total_2015<=200)
hi_cut_plus_LA <- hi_cut %>% filter(Med_HH_Income_Percent_of_State_Total_2015>200)

point_income_per_state_avg <- ggplot() +
  geom_smooth(data=hi_cut_minus_LA, aes(Median_Household_Income_2015, Med_HH_Income_Percent_of_State_Total_2015)) + 
  geom_point(data=hi_cut_minus_LA, aes(Median_Household_Income_2015, Med_HH_Income_Percent_of_State_Total_2015, color=factor(State)),size=2) +
  geom_point(data=hi_cut_plus_LA, aes(Median_Household_Income_2015, Med_HH_Income_Percent_of_State_Total_2015), color="black",size=3) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_continuous(labels = function(x) paste0("$", scales::comma(x))) +
  labs(x='Median household income', y="Household income as percent of state average", color="State",
       title="Census areas with the highest household incomes are almost all located in states with high\naverage incomes.", 
       subtitle="\nTwo major exceptions are Williamson County, TN, a suburb of Nashville that boasts many of Tennessee's\nrichest, and Los Alamos County, NM, home to a well-paying national laboratory... and not much else.") +
    geom_label_repel(data=hi_cut_plus_LA, aes(Median_Household_Income_2015, Med_HH_Income_Percent_of_State_Total_2015, label = Area_name),
                  box.padding   = 0.35, 
                  point.padding = 0.85,
                  segment.color = 'grey50') +
  scale_color_brewer(palette = "Set1")

point_income_per_state_avg

ggsave(filename="point_income_per_state_avg.pdf", device = "pdf", plot = point_income_per_state_avg, height = 5, width = 8, units = "in")
  
```


```{r}

ggplot(data=hi_cut, aes(Percent.of.adults.with.a.bachelor.s.degree.or.higher..2011.2015, Unemployment_rate_2015))+geom_point()+
    geom_label_repel(aes(label = Area_name),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')

#aes(reorder(sources, deaths), deaths, fill=deaths)
```

```{r}
boring <- ggplot() + geom_point(data = census, aes(Median_Household_Income_2015, Unemployment_rate_2015, color=Percent.of.adults.with.a.bachelor.s.degree.or.higher..2011.2015, label=State)) +
  labs(y = "Unemployment rate", x = "Median household income", title = "Census areas where a majority of the population holds a bachelors degree\nor higher are among the highest paid and most likley to be employed", color="Percent of population with \na bachelors or higher") +
  scale_color_viridis(option="plasma", limits=c(0,80), breaks=seq(0,80,20), labels=function(x) paste0(x, "%")) +
  coord_trans(x="log", y="log") +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_continuous(labels = function(x) paste0("$", scales::comma(x)))
boring

ggsave(filename="bachelors_employed_paid.pdf", device = "pdf", plot = boring, height = 5, width = 8, units = "in")

ggplot(data=states,aes(Median_Household_Income_2015, Unemployment_rate_2015)) + geom_point(aes(color=Percent.of.adults.with.a.bachelor.s.degree.or.higher..2011.2015)) +
  labs(y = "Unemployment rate", x = "Med HH income", title = "Census sandbox", color="BS+") +
    geom_label_repel(aes(label = State),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')
  
```

```{r}
# find difference in percent with bachelors plus and check HH income
bachelors_increase <- counties %>% mutate(bs_increase = Percent.of.adults.with.a.bachelor.s.degree.or.higher..2011.2015 - Percent.of.adults.with.a.bachelor.s.degree.or.higher..1990)

bs_decrease <- bachelors_increase %>% filter(bs_increase <= 0 & Median_Household_Income_2015>60000)
bs_decrease

ggplot(data=bachelors_increase, aes(x=bs_increase, y=Median_Household_Income_2015)) + geom_point()

ggplot(data=bs_decrease, aes(x=bs_increase, y=Median_Household_Income_2015)) + geom_point(aes(color=factor(State))) +
  geom_label_repel(aes(label = Area_name),
                  box.padding   = 1, 
                  point.padding = 1.9,
                  segment.color = 'grey50')
```


```{r}
wi <- census %>% filter(State=="WI" & Area.name != "Wisconsin") %>% mutate(pop_gain = R_birth_2015-R_death_2015) %>%
  select(c(X2013.Urban.Influence.Code, R_NET_MIG_2015, Median_Household_Income_2015, Unemployment_rate_2015, pop_gain)) %>% group_by(X2013.Urban.Influence.Code) %>% summarise(mean(Unemployment_rate_2015), mean(pop_gain)) %>% rename(unemp = "mean(Unemployment_rate_2015)", pop_gain = "mean(pop_gain)") 
View(wi)

mi <- census %>% filter(State=="MI" & Area.name != "Michigan") %>% select(c(X2013.Urban.Influence.Code, R_NET_MIG_2015, Unemployment_rate_2015)) %>% group_by(X2013.Urban.Influence.Code) %>% summarise(mean(Unemployment_rate_2015))
View(mi)

nm <- census %>% filter(State=="NM" & Area.name != "New Mexcio") %>% select(c(X2013.Urban.Influence.Code, R_NET_MIG_2015, Unemployment_rate_2015)) %>% group_by(X2013.Urban.Influence.Code) %>% summarise(mean(Unemployment_rate_2015))
View(nm)

ca <- census %>% filter(State=="CA" & Area.name != "California") %>% select(c(X2013.Urban.Influence.Code, R_NET_MIG_2015, Unemployment_rate_2015)) %>% group_by(X2013.Urban.Influence.Code) %>% summarise(mean(Unemployment_rate_2015))
View(ca)
```


```{r}
wi_unemploy_nat_pop_change <- ggplot(data=wi) + geom_bar(aes(reorder(X2013.Urban.Influence.Code, unemp), unemp, fill=pop_gain), stat="identity") +
  labs(x="Urban influence (1 most urban to 12 least urban)", y="Unemployment rate", fill="Natural population change\n(birth rate - death rate)", 
       title="Rural counties in Wisconsin have both a higher unemployment rate\nand a lower rate of natural population increase than more urbanized counties") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_fill_viridis(option="cividis")
wi_unemploy_nat_pop_change

ggsave(filename="wi_unemploy_nat_pop_change.pdf", device = "pdf", plot = wi_unemploy_nat_pop_change, height = 5, width = 8, units = "in")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 